The Atala PRISM Protocol

### 2022-12-16 prism-backend/research/prism-protocol/did_atala.tex
Rewrote ideal functionality using global subroutines.
Commiting before applying more changes. Forgot what I did here...

\begin{figure}
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Upon receiving a call $\IdealGPKIDID^{\P}(\cmd,\sid,\arg)$
          from some party $P$ to run command \cmd, in session \sid, with arguments
          \arg, $\IdealGPKIDID^{\P}$ operates as follows.}
        \textrm{If the received \cmd and \arg pair does not meet any of
          the following options, the request is ignored.}
        \textrm{\P is a boolean predicate defining concrete conditions for
          acceptable DID creations, updates and deactivations.}
        \textrm{\didOP is a list of submitted (yet, possibly unconfirmed) DID
          operations, and \DID is the authoritative state of published DIDs.}
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \setind{(\lbl_i,\typ_i)}{i\in[n]}$}{ 
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\cdot) \in \DID \lor
          \P(\setind{(\lbl_i,\typ_i)}{i\in[n]}) = 0: \\
          \pcind \pcreturn \bot \\
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{CreateOk}, \sid, \did,
          \sval=\setind{\val_i}{i\in[n]}) \\
          \hspace*{2.7em}\text{from}~\Sim \\          
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \didOp \gets (\did,\setind{(\lbl_i,\typ_i,\val_i)}{i\in[n]}) \\
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit}, \didOp)~\text{to}~\IdealGdledger \\
          \ucio{Output}~(\uccmd{ProcessedC},\sid,\did,\sval)
          % \pcind \DID \gets \DID \cup \lbrace (P,\did,\lbrace
          % (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          % \ucio{Output}~(\uccmd{Created},\sid,\did,\sval,t)
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = \did$}{
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\sval) \notin \DID \lor \P(\sval) = 0: \\
          \pcind \pcreturn \bot \\
          \ucio{Output}~(\uccmd{ReadOk},\sid,\did,\sval)~\text{to}~P
        }    
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        % \arg = (did, \sval= \lbrace (\lbl_i,\typ_i) \rbrace_{i\in[n]})\$}{
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d,\sval_a)$}{
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\sval_0) \notin \DID: \pcreturn \bot \\
          \pcif \P((\sval_0 \setminus \sval_d) \cup \sval_a) = 0:
          \pcreturn \bot \\
          \ucsend~(\uccmd{Update}, \sid,\did,\sval_d,\sval_a)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{UpdOk},\sid,\did,\lbrace \val^a_i \rbrace_{i\in[n^a]})~
          \text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \lbrace (\lbl_i^a,\typ_i^a) \rbrace_{i\in[n^a]} \gets \sval_a \\
          %\lbrace \lbl_i^d \rbrace_{i\in[n^d]} \gets \sval_d \\          
          \sval_1 \gets (\sval_0 \setminus \sval_d) \cup
          \setind{(\lbl^a_i,\typ^a_i,\val^a_i)}{i\in[n^a]} \\
          \didOp \gets (\did,\sval_1);~
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit}, \didOp)~\text{to}~\IdealGdledger \\
          \ucio{Output}~(\uccmd{ProcessedU},\sid,\did,\didOp)
        }
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace \lbl_i \rbrace_{i\in[n]})$}{
          \DID \gets \DIDReadLdg \\
          \pcif (P,did,\sval_0) \notin \DID: \pcreturn \bot \\
          \sval_1 \gets \sval_0 \setminus \lbrace \lbl_i \rbrace_{i\in[n]} \\
          \pcif \P(\sval_1) = 0: \pcreturn \bot \\
          \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\          
          \didOp \gets (\did,\sval_1);
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit},\didOp)~\text{to}~\IdealGdledger \\          
          \ucio{Output}~(\uccmd{ProcessedD},\sid,\did,\didOp)
        }
      \end{minipage}
    }
    \vspace*{0.25em} \\
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \centering
        \procedure[linenumbering]{$\DIDReadLdg$}{
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGdledger~\text{and recv}~
          (\uccmd{ReadOk},\DID)~\text{back} \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock ~\text{and recv}~
          (\uccmd{ReadOk},\tau)~\text{back} \\
          \pcif \exists (\cdot,\tau',\didOp) \in \didOP~\suchthat~
          |\tau - \tau'| > \delta \land \P(\didOp) = 1 \land \didOp \notin \DID:
          \pcreturn \bot \\
          \pcreturn \DID
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealGPKIDID. We abuse set
    notation, and write $\lbrace (a_i,\dots) \rbrace_{i\in[n]} \setminus
    \lbrace (b_j,\dots) \rbrace_{j\in[m]}$ to mean subtracting from the first
    set the elements present in the second set, whenever $a_i = b_j$ for some
    $i,j$, even if the size of the $(a_i,\dots)$ and $(b_j,\dots)$ tuples do not
    match.}
  \label{fig:fpkidid}
\end{figure}

>>>>

\begin{figure}
  \begin{framed}
    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \textrm{Upon receiving a call $\IdealGPKIDID^{\P}(\cmd,\sid,\arg)$
          from some party $P$ to run command \cmd, in session \sid, with arguments
          \arg, $\IdealGPKIDID^{\P}$ operates as follows.}
        \textrm{If the received \cmd and \arg pair does not meet any of
          the following options, the request is ignored.}
        \textrm{\P is a boolean predicate defining concrete conditions for
          acceptable DID creations, updates and deactivations.}
        \textrm{\didOP is a list of submitted (yet, possibly unconfirmed) DID
          operations, and \DID is the authoritative state of published DIDs.}
        \todo{Think about what happens if two conflicting ops (esp. create)
          are submitted with diff $t < \delta$.}
        \todo{Idea: func keeps track of DID ownership, associated to parties.
          Also, replace \didOP with an UTXO, that gets updated like the delay
          ledger (i.e., upon applying a new op, remove it from the set.)}
      \end{minipage}
    }
    \vspace*{0.5em}

    \scalebox{0.9}{
      \begin{minipage}[t]{0.55\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Create} \land \arg =
          \setind{(\lbl_i,\typ_i)}{i\in[n]}$}{ 
          %\DID \gets \DIDReadLdg \\          
          \ucsend~(\uccmd{Create}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{CreateOk}, \sid, \did,
          \sval=\setind{\val_i}{i\in[n]}) \\
          \hspace*{2.7em}\text{from}~\Sim \\
          % \pcif (\did,\cdot) \in \DID \lor
          % \P(\setind{(\lbl_i,\typ_i)}{i\in[n]}) = 0: \\
          % \pcind \pcreturn \bot \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \didOp \gets (\did,\setind{(\lbl_i,\typ_i,\val_i)}{i\in[n]}) \\
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit}, \didOp)~\text{to}~\IdealGdledger \\
          \ucio{Output}~(\uccmd{ProcessedC},\sid,\did,\sval)
          % \pcind \DID \gets \DID \cup \lbrace (P,\did,\lbrace
          % (\lbl_i,\typ_i,\val_i) \rbrace_{i\in[n]}) \rbrace \\
          % \ucio{Output}~(\uccmd{Created},\sid,\did,\sval,t)
        }
        \vspace*{0.25em} \\

        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Update} \land
          \arg = (did, \sval_d,\sval_a)$}{
          \DID \gets \DIDReadLdg \\
          \pcif (P,\did) \notin \didOWN: \pcreturn \bot \\
          \pcif (\did,\sval_0) \notin \DID: \pcreturn \bot \\ %\lor (P,\cdot,(\did,\cdot))
          % \notin \didOP: \\
          % \pcind \pcreturn \bot \\
          % \pcif \P((\sval_0 \setminus \sval_d) \cup \sval_a) = 0:
          % \pcreturn \bot \\
          \ucsend~(\uccmd{Update}, \sid,\did,\sval_d,\sval_a)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{UpdOk},\sid,\did,\setind{\val^a_i}{i\in[n^a]})~
          \text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\
          \lbrace (\lbl_i^a,\typ_i^a) \rbrace_{i\in[n^a]} \gets \sval_a \\
          % \lbrace \lbl_i^d \rbrace_{i\in[n^d]} \gets \sval_d \\          
          \sval_1 \gets (\sval_0 \setminus \sval_d) \cup
          \setind{(\lbl^a_i,\typ^a_i,\val^a_i)}{i\in[n^a]} \\
          \didOp \gets (\did,\sval_1);~
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit}, \didOp)~\text{to}~\IdealGdledger \\
          \ucio{Output}~(\uccmd{ProcessedU},\sid,\did,\didOp)
        }        
      \end{minipage}
    }
    \scalebox{0.9}{
      \begin{minipage}[t]{0.5\textwidth}
        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Read} \land
          \arg = \did$}{
          \DID \gets \DIDReadLdg \\
          \pcif (\did,\sval) \notin \DID: \pcreturn \bot \\ % \lor \P(\sval) = 0: \\
          %\pcind \pcreturn \bot \\
          \ucio{Output}~(\uccmd{ReadOk},\sid,\did,\sval)~\text{to}~P
        }
        \vspace*{0.25em} \\

        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Deact} \land \arg =
          (did, \lbrace \lbl_i \rbrace_{i\in[n]})$}{
          \DID \gets \DIDReadLdg \\
          \pcif (P,\did) \notin \didOWN: \pcreturn \bot \\                    
          \pcif (P,did,\sval_0) \notin \DID: \pcreturn \bot \\ % \lor (P,\cdot,(\did,\cdot))
          % \notin \didOP: \\
          % \pcind \pcreturn \bot \\
          \sval_1 \gets \sval_0 \setminus \lbrace \lbl_i \rbrace_{i\in[n]} \\
          % \pcif \P(\sval_1) = 0: \pcreturn \bot \\
          \ucsend~(\uccmd{Deact}, \sid, \arg)~\text{to}~\Sim \\
          \ucrecv~(\uccmd{DeactOk},\sid)~\text{from}~\Sim \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock \\
          \ucrecv~(\uccmd{ReadOk},\tau)~\text{from}~\IdealGclock \\          
          \didOp \gets (\did,\sval_1);
          \didOP \gets \didOP \cup \set{(P,\tau,\didOp)} \\
          \ucsend~(\uccmd{Submit},\didOp)~\text{to}~\IdealGdledger \\          
          \ucio{Output}~(\uccmd{ProcessedD},\sid,\did,\didOp)
        }
%        \vspace*{0.05em} \\

        \procedure[linenumbering]{$\pcif \cmd = \uccmd{Extend} \land \arg =
          \didOP'$}{
          \pcif P \neq \Sim \lor \didOP' \subseteq \didOP: \pcreturn \bot \\
          \DID \gets \DIDReadLdg \\          
          %\lor~\\
          %\pcind \exists \didOp \in \didOP'~\suchthat~\P(\didOp) = 0: \\
          %\pcind \pcreturn \bot \\
          \pcfor \didOp = (P,\cdot,(\did,\cdot)) \in \didOP': \\
          \pcind \pcif \P(\didOp,\DID) = 0: \pcreturn \bot \\
          \pcind \didOWN \gets (P,\did) \\
          \didOP \gets \didOP \setminus \didOP' \\
          \ucsend~(\uccmd{Extend},\didOP')~\text{to}~\IdealGledger
        }
      \end{minipage}
    }
    \vspace*{0.25em} \\

    \scalebox{0.9}{
      \begin{minipage}[t]{\textwidth}
        \centering
        \procedure[linenumbering]{$\DIDReadLdg$}{
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGdledger~\text{and recv}~
          (\uccmd{ReadOk},\DID)~\text{back} \\
          \ucsend~(\uccmd{Read})~\text{to}~\IdealGclock ~\text{and recv}~
          (\uccmd{ReadOk},\tau)~\text{back} \\
          \pcif \exists (\cdot,\tau',\didOp) \in \didOP~\suchthat~
          |\tau - \tau'| > \delta \land \P(\didOp) = 1 \land \didOp \notin \DID:
          \pcreturn \bot \\
          \pcreturn \DID
        }
      \end{minipage}
    }
  \end{framed}
  \caption{Ideal DID-based PKI functionality, \IdealGPKIDID. We abuse set
    notation, and write $\lbrace (a_i,\dots) \rbrace_{i\in[n]} \setminus
    \lbrace (b_j,\dots) \rbrace_{j\in[m]}$ to mean subtracting from the first
    set the elements present in the second set, whenever $a_i = b_j$ for some
    $i,j$, even if the size of the $(a_i,\dots)$ and $(b_j,\dots)$ tuples do not
    match.}
  \label{fig:fpkidid}
\end{figure}
